The role of cognateness in native spoken word recognition

### 2024-10-16 manuscript/manuscript.tex
Fix figures

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{tipa}
\usepackage{booktabs}
\DeclareTextFontCommand\textphonetic{\ipafont}
\usepackage[utf8]{inputenc}
\usepackage{CJKutf8}
\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

When some German speakers listened to the song \emph{The Power} by SNAP!
(\emph{Coyote Ugly}, 2000), many of them misheard the line ``I've got
the power'' as \emph{Agatha Bauer!} When Michael Jackson vocalised
``ma-ma-coo-sah'' in his song \emph{Wanna Be Startin' Somethin'}
(\emph{Thriller}, 1982), some Dutch speakers understood \emph{mama
appelsap} {[}mama applejuice{]}. When Spanish speakers listen to the
line ``Circumvent your thick ego'' from the song \emph{Pictures} by
System of a Down (\emph{Steal this Album!}, 2002), they tend to hear
\emph{Sácame de aquí, King-Kong} {[}take me out of here, King-Kong{]}.
When Japanese speakers listen to the line ``I want to hold your hand''
in the homonym song by The Beatles (\emph{Meet The Beatles}, 1964), they
often mishear \emph{Aho na hōnyōhan} {[}stupid urinator{]}. Outrageous
as these examples might sound to speakers of other languages, many
readers may know a few cases in their own native language. This auditory
illusion can feel quite real, and often inevitable after the first time
it happens. In Japanese, this phenomenon takes the name \emph{Soramimi}
(lit. ``empty ear''), after the \emph{Soramimi Hour}---a section of the
program \emph{Tamori Club} hosted by comedian Morita Kazuyoshi
(Tamori)---in which instances of this illusion were presented with
comedy purposes. Soramimi is a particular case of \emph{homophonic
translation}: words or phrases in one language are translated into
similar-sounding words and phrases in a different language, without
necessarily preserving the meaning. This phenomenon extends beyond song
lyrics, and may occur anytime when listening to non-native speech. For
instance, when listening to ``You know me?'' (\textipa{/ju: n@U mi:}/)
in English, Japanese-native listeners tend to understand \emph{yunomi}
(/\textipa{jUnomi}/), which is a type of Japanese teacup. Overall, the
occurrence of Soramimi---and more generally, homophonic
translations---suggests that speech in an unfamiliar language has the
potential to activate lexical representations in the native language.

@@

After data collection, participants' answers were manually coded into
the following categories: \emph{Correct}, \emph{Typo}, \emph{Wrong},
\emph{False friend}, \emph{Other}. A response was coded as
\emph{Correct} if the provided string of characters was identical to the
orthographic form of the correct translation. A response was coded as
\emph{Typo} if the participant provided a string of characters only one
edit distance (addition, deletion, or substitution) apart from the
orthographic form of the correct translation (e.g., ``pengiun'' instead
of ``penguin''), as long as the response did not correspond to a
distinct English word. A response was coded as \emph{False friend} if
the participant provided a phonologically similar incorrect translation.
Responses not meeting the criteria for previous categories were labelled
as \emph{Wrong} or \emph{Other} (see Data analysis section for more
details). Both \emph{Correct} and \emph{Typo} responses were considered
as correct, while \emph{Wrong} and \emph{False friend} responses were
considered as incorrect. \emph{Other} responses were excluded from data
analysis. Trials in which participants took longer than 10 seconds to
respond were also excluded. Participants contributed a total of 9152
valid trials (5,694 in Catalan, 3,458 in Spanish). The task took
approximately 15 minutes to be completed.

@@

\[
\begin{aligned}
&\textbf{Likelihood}  \\
y_{i} \sim & \text{Bernoulli}(p_{i}) \\ \\
&\textbf{Parameters}  \\
\text{Logit}(p_{i}) = &  \beta_{0[p,w]} + \beta_{1[p]} \text{Frequency}_{i} + \beta_{2[p]} \text{PTHN}_i + \beta_{3[p]} \text{Similarity}_i + \beta_{4[p]} (\text{PTHN}_i \times \text{Similarity}_i) \\
\beta_{0-6[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
\beta_{1-6[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
&\textbf{Prior}  \\
\mu_{\beta_{p,w}}  \sim &  \mathcal{N}(0, 0.1) \\
\sigma_{\beta_{p}},  \sigma_{\beta_{w}} \sim & \text{HalfCauchy}(0, 0.1) \\
\rho_{p}, \rho_{w} \sim & \text{LKJCorr}(8) \\
\end{aligned}
\] ::: \{eq-1\}

To test the practical relevance of each predictor we followed Kruschke
\& Liddell (\citeproc{ref-kruschke2018bayesian}{2018}). We first
specified a region of practical equivalence (\emph{ROPE}) around zero
({[}-0.1, +0.1{]}, in the logit scale). This area indicates the values
of the regression coefficients that we considered as equivalent to zero.
We then computed the 95\% posterior credible intervals (CrI) of the
regression coefficients of interest. Finally, we calculated the
proportion of the 95\% CrI that fell inside the ROPE. This proportion
indicates the probability that the true value of the coefficient is
equivalent to zero. All analyses were performed in R environment
(\citeproc{ref-rcoreteam2013language}{R Core Team, 2013}). We used the
tidyverse family of R packages
(\citeproc{ref-wickham2019welcome}{Wickham et al., 2019}) to process
data and to generate figures. We used the \texttt{brms} R package
(\citeproc{ref-burkner2017brms}{Bürkner, 2017}) using the
\texttt{cmdstanr} back-end to the Stan probabilistic language
(\citeproc{ref-carpenter2017stan}{Carpenter et al., 2017}) to estimate
and compare the models (see Appendix 1 for mode details on the models).

@@

We collected data for a total of 6,446 trials completed by 72 unique
participants. Of those trials, 2,988 were provided by 36 unique
participants who listened to Catalan words, and 3,458 trials were
provided by 36 unique participants who listened to Spanish words. We
excluded trials in which participants did not enter any text (\emph{n} =
72), in which a response in a language other than English was provided
(e.g., \texttt{agua}, \emph{n} = 51), in which participants did not
provide a whole word (e.g., \texttt{f}, \emph{n} = 5), and in which
participants added comments to the experimenter (e.g., \texttt{unsure},
\emph{n} = 13). In addition, we excluded data from participants that
self-rated their oral and/or written skills in Catalan and Spanish, or
any other second language as four or higher in a five-point scale
(\emph{n} = 2), were diagnosed with a language (\emph{n} = 1), or did
not contribute more than 80\% of valid trials (\emph{n} = 9).

@@

Table~\ref{tbl-dataset} shows a summary of participants' accuracy across
Experiments 1, 2, and 3. MCMC chains in the model showed strong evidence
of convergence (\(\hat{R}<1.01\)) (see Appendix 2 for more detailed
model diagnostics). Participants translating Catalan words and
participants translating Spanish words performed equivalently, as
indicated by the regression coefficient of \emph{Group} (\(\beta\) =
-0.199, 95\% CrI = {[}-0.512, 0.129{]}, \emph{p}(ROPE) = 0.238).
Overall, participants responded less accurately to words with more CLPNs
than to words with fewer CLPNs, regardless of the amount of phonological
similarity between the presented word and its translation. This is
indicated by the size of the regression coefficient of the two-way
interaction between \emph{Similarity} and \emph{CLPN} (\(\beta\) =
-0.653, 95\% CrI = {[}-0.973, -0.313{]}, \emph{p}(ROPE) = 0). As
anticipated, participants' performance benefited from an increase in
\emph{Similarity} (\(\beta\) = 7.133, 95\% CrI = {[}6.606, 7.735{]},
\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
effect (\(\beta\) = 0.009, 95\% CrI = {[}-0.11, 0.111{]}, \emph{p}(ROPE)
= 0.925). Figure~\ref{fig-epreds-1} illustrates the posterior of the
average predictions of the model for words with different values of
\emph{Similarity} and \emph{CLPN}. Figure~\ref{fig-coefs} shows a
graphic summary of the posterior distribution of the regression
coefficients of interest.

@@

We collected data for a total of 5,412 trials completed by 33 unique
participants. We excluded trials in which participants did not enter any
text (\emph{n} = 44), in which a response in a language other than
English was provided (e.g., \texttt{agua}, \emph{n} = 51), in which
participants did not provide a whole word (e.g., \texttt{f}, \emph{n} =
7), and in which participants added comments to the experimenter (e.g.,
\texttt{unsure}, (\emph{n} = 1). In addition, we excluded data from
participants that self-rated their oral and/or written skills in Catalan
and Spanish, or any other second language as four or higher in a
five-point scale (\emph{n} = 22), were diagnosed with a developmental
language disorder (\emph{n} = 1), or did not contribute more than 80\%
of valid trials (\emph{n} = 9). After applying trial-level and
participant-level inclusion criteria, the resulting dataset included
1,662 trials provided by 42 unique participants. Of those trials, 1,662
were provided by 21 unique participants who listened to Catalan words,
and 1,662 trials were provided by 21 unique participants who listened to
Spanish words. Responses given by participants were 5.6 characters long
on average (\emph{Median} = 5, \emph{SD} = 1.6, Range = 2-12).

@@

\begin{longtable}{l|rr}

@@

\toprule
\multicolumn{1}{l}{} & Accuracy (\%) & SE \\ 
\midrule\addlinespace[2.5pt]
\multicolumn{3}{l}{Experiment 1 - cat-ENG} \\ 
\midrule\addlinespace[2.5pt]
cavall /kəβaʎ/ - horse /hɔs/ & $17.14$ & $6.37$ \\ 
llibre /ʎiβɾə/ - book /bʊk/ & $17.14$ & $6.37$ \\ 
camisa /kəmizə/ - shirt /ʃɜt/ & $16.67$ & $6.21$ \\ 
poma /poma/ - apple /æpl/ & $16.67$ & $6.21$ \\ 
cama /kamə/ - leg /lɛg/ & $11.11$ & $5.24$ \\ 
\midrule\addlinespace[2.5pt]
\multicolumn{3}{l}{Experiment 1 - spa-ENG} \\ 
\midrule\addlinespace[2.5pt]
pantalón /paŋtalon/ - trousers /traʊzəz/ & $77.42$ & $7.51$ \\ 
naranja /naɾaŋxa/ - orange /ɒrɪnʤ/ & $41.94$ & $8.86$ \\ 
leche /leʧe/ - milk /mɪlk/ & $35.48$ & $8.59$ \\ 
toro /toɾo/ - bull /bʊl/ & $33.33$ & $8.61$ \\ 
libro /liβɾo/ - book /bʊk/ & $30.00$ & $8.37$ \\ 
cebra /θebɾa/ - zebra /zibrə/ & $29.03$ & $8.15$ \\ 
pan /pan/ - bread /brɛd/ & $29.03$ & $8.15$ \\ 
pollo /poʎo/ - chicken /ʧɪkɪn/ & $26.67$ & $8.07$ \\ 
jirafa /xiɾafa/ - giraffe /ʤɪrɑf/ & $20.69$ & $7.52$ \\ 
perro /pero/ - dog /dɒɡ/ & $16.13$ & $6.61$ \\ 
pluma /pluma/ - feather /fɛðə/ & $16.13$ & $6.61$ \\ 
puerta /pwerta/ - door /dɔ/ & $16.13$ & $6.61$ \\ 
pie /pje/ - foot /fʊt/ & $12.90$ & $6.02$ \\ 
caballo /kaβaʎo/ - horse /hɔs/ & $10.34$ & $5.66$ \\ 
bocadillo /bokadiʎo/ - sandwich /sænwɪʤ/ & $10.00$ & $5.48$ \\ 
globo /ɡloβo/ - balloon /bəlun/ & $10.00$ & $5.48$ \\ 
\midrule\addlinespace[2.5pt]
\multicolumn{3}{l}{Experiment 2 - cat-SPA} \\ 
\midrule\addlinespace[2.5pt]
fulla /fuʎə/ - hoja /oxa/ & $30.43$ & $9.59$ \\ 
ull /uʎ/ - ojo /oxo/ & $21.74$ & $8.60$ \\ 
got /gɔt/ - vaso /baso/ & $20.00$ & $8.00$ \\ 
entrepà /entɾəpa/ - bocadillo /bokadiʎo/ & $13.04$ & $7.02$ \\ 
mirall /miɾaʎ/ - espejo /espexo/ & $12.50$ & $6.75$ \\ 
\bottomrule

@@

{[}ADD HERE LEXICAL FREQUENCY OF CATALAN AND SPANISH WORDS IN ENGLISH,
AND OF CATALAN WORDS IN SPANISH{]}

@@

We collected data for a total of 6,016 trials completed by 64 unique
participants. Of those trials, 2,752 were provided by 32 unique
participants who listened to Catalan words, and 3,264 trials were
provided by 32 unique participants who listened to Spanish words. We
excluded trials in which participants did not enter any text (\emph{n} =
), in which a response in a language other than English was provided
(e.g., \texttt{agua}, \emph{n} = ), in which participants did not
provide a whole word (e.g., \texttt{f}, \emph{n} = ), and in which
participants added comments to the experimenter (e.g., \texttt{unsure},
(\emph{n} = ). In addition, we excluded data from participants that
self-rated their oral and/or written skills in Catalan and Spanish, or
any other second language as four or higher in a five-point scale
(\emph{n} = 2), were diagnosed with a language (\emph{n} = 0), or did
not contribute more than 80\% of valid trials (\emph{n} = 1).

@@

MCMC chains in the model showed strong evidence of convergence
(\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
Overall, participants reported prior knowledge more often for that
Spanish words with unexpectedly high accuracy (see Discussion in
Experiment 2) than for words with expected accuracy (see
Figure~\ref{fig-knowledge}). Participants reported prior knowledge of
Catalan words with unexpected accuracy as often as those with expected
accuracy. This suggests that participants in Experiment 1 may have
relied, to some extent, on their prior knowledge about form-meaning
mappings to correctly translate some Spanish words. To isolate such an
effect of prior Spanish knowledge, we run the same analysis as in
Experiment 1 on the newly collected translations from Experiment 3, now
excluding responses to words in which participants reported prior
knowledge.

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-knowledge-1.pdf}}

\caption{\label{fig-knowledge}Catalan/Spanish prior word knowledge as
reported by English native participants in Experiment 3. (A) Average
proportion of participants that reported prior knowledge for words with
surprisingly high accuracy (no phonological similarity with the correct
translation, accuracy higher than 10\%), and for words with expected
accuracy (low similarity-low accuracy, or high similarity-high
accuracy). (B) Average accuracy for words with expected and unexpected
accuracy.}

@@

\section{Pooled analyses of Experiments 1 and
3}\label{pooled-analyses-of-experiments-1-and-3}

Across Experiments 1 and 3, we found strong evidence that participants
efficiently exploited phonological similarity to provide accurate
translations for words in an unfamiliar language, provided that few
phonological neighbours of higher lexical frequency were present.
Figure~\ref{fig-coefs} summarizes the posterior distribution of the
regression coefficients of the models in Experiments 1 to 3.

\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-coefs-1.pdf}}

\caption{\label{fig-coefs}}

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-best2001discrimination}
Best, C. T., McRoberts, G. W., \& Goodell, E. (2001). Discrimination of
non-native consonant contrasts varying in perceptual assimilation to the
listener's native phonological system. \emph{The Journal of the
Acoustical Society of America}, \emph{109}(2), 775--794.

@@

@@

@@

>>>>

\usepackage{booktabs}
\usepackage{caption}
\usepackage{longtable}
\usepackage{colortbl}
\usepackage{array}
\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{rotating}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\usepackage{tipa}
\usepackage{booktabs}
\DeclareTextFontCommand\textphonetic{\ipafont}
\usepackage[utf8]{inputenc}
\usepackage{CJKutf8}
\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

When some German speakers listened to the song \emph{The Power} by SNAP!
(\emph{Coyote Ugly}, 2000), many of them misheard the line ``I've got
the power'' as \emph{Agatha Bauer!} When Michael Jackson vocalised
``ma-ma-coo-sah'' in his song \emph{Wanna Be Startin' Somethin'}
(\emph{Thriller}, 1982), some Dutch speakers understood \emph{mama
appelsap} {[}mama applejuice{]}. When Spanish speakers listen to the
line ``Circumvent your thick ego'' from the song \emph{Pictures} by
System of a Down (\emph{Steal this Album!}, 2002), they tend to hear
\emph{Sácame de aquí, King-Kong} {[}take me out of here, King-Kong{]}.
When Japanese speakers listen to the line ``I want to hold your hand''
in the homonym song by The Beatles (\emph{Meet The Beatles}, 1964), they
often mishear \emph{Aho na hōnyōhan} {[}stupid urinator{]}. Outrageous
as these examples might sound to speakers of other languages, many
readers may know a few cases in their own native language. This auditory
illusion can feel quite real, and often inevitable after the first time
it happens. In Japanese, this phenomenon takes the name \emph{Soramimi}
(lit. ``empty ear''), after the \emph{Soramimi Hour}---a section of the
program \emph{Tamori Club} hosted by comedian Morita Kazuyoshi
(Tamori)---in which instances of this illusion were presented with
comedy purposes. Soramimi is a particular case of \emph{homophonic
translation}: words or phrases in one language are translated into
similar-sounding words and phrases in a different language, without
necessarily preserving the meaning
(\citeproc{ref-gasparov2006semen}{Gasparov, 2006}). This phenomenon
extends beyond song lyrics, and may occur anytime when listening to
non-native speech. For instance, when listening to ``You know me?''
(\textipa{/ju: n@U mi:}/) in English, Japanese-native listeners tend to
understand \emph{yunomi} (/\textipa{jUnomi}/), which is a type of
Japanese teacup. Overall, the occurrence of Soramimi---and more
generally, homophonic translations---suggests that speech in an
unfamiliar language has the potential to activate lexical
representations in the native language.

@@

After data collection, participants' answers were manually coded into
the following categories: \emph{Correct}, \emph{Typo}, \emph{Wrong},
\emph{False friend}, \emph{Other}. A response was coded as
\emph{Correct} if the provided string of characters was identical to the
orthographic form of the correct translation. A response was coded as
\emph{Typo} if the participant provided a string of characters only one
edit distance (addition, deletion, or substitution) apart from the
orthographic form of the correct translation (e.g., ``pengiun'' instead
of ``penguin''), as long as the response did not correspond to a
distinct English word. A response was coded as \emph{False friend} if
the participant provided a phonologically similar incorrect translation.
Responses not meeting the criteria for previous categories were labelled
as \emph{Wrong} or \emph{Other} (see Data analysis section for more
details). Both \emph{Correct} and \emph{Typo} responses were considered
as correct, while \emph{Wrong} and \emph{False friend} responses were
considered as incorrect. \emph{Other} responses were excluded from data
analysis. Trials in which participants took longer than 10 seconds to
respond were also excluded. Participants contributed a total of 9,152
valid trials (5,694 in Catalan, 3,458 in Spanish). The task took
approximately 15 minutes to be completed.

@@

\begin{equation}\phantomsection\label{eq-1}{
\begin{aligned}
&\textbf{Likelihood}  \\
y_{i} \sim & \text{Bernoulli}(p_{i}) \\ \\
&\textbf{Parameters}  \\
\text{Logit}(p_{i}) = &  \beta_{0[p,w]} + \beta_{1[p]} \text{Frequency}_{i} + \beta_{2[p]} \text{PTHN}_i + \beta_{3[p]} \text{Similarity}_i + \beta_{4[p]} (\text{PTHN}_i \times \text{Similarity}_i) \\
\beta_{0-6[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
\beta_{1-6[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
&\textbf{Prior}  \\
\mu_{\beta_{p,w}}  \sim &  \mathcal{N}(0, 0.1) \\
\sigma_{\beta_{p}},  \sigma_{\beta_{w}} \sim & \text{HalfCauchy}(0, 0.1) \\
\rho_{p}, \rho_{w} \sim & \text{LKJCorr}(8) \\
\end{aligned}
}\end{equation}

To test the practical relevance of each predictor we followed Kruschke
\& Liddell (\citeproc{ref-kruschke2018bayesian}{2018}). We first
specified a region of practical equivalence (ROPE) around zero ({[}-0.1,
+0.1{]}, in the logit scale). This area indicates the values of the
regression coefficients that we considered as equivalent to zero. We
then computed the 95\% posterior credible intervals (CrI) of the
regression coefficients of interest. Finally, we calculated the
proportion of the 95\% CrI that fell inside the ROPE. This proportion
indicates the probability that the true value of the coefficient is
equivalent to zero. All analyses were performed in R environment
(\citeproc{ref-rcoreteam2013language}{R Core Team, 2013}). We used the
tidyverse family of R packages
(\citeproc{ref-wickham2019welcome}{Wickham et al., 2019}) to process
data and to generate figures. We used the \texttt{brms} R package
(\citeproc{ref-burkner2017brms}{Bürkner, 2017}) using the
\texttt{cmdstanr} back-end to the Stan probabilistic language
(\citeproc{ref-carpenter2017stan}{Carpenter et al., 2017}) to estimate
and compare the models (see Appendix 1 for mode details on the models).

@@

We collected data for a total of 6,446 trials completed by 72 unique
participants. Of those trials, 2,988 were provided by 36 unique
participants who listened to Catalan words, and 3,458 trials were
provided by 36 unique participants who listened to Spanish words. We
excluded trials in which participants did not enter any text (\emph{n} =
72), in which a response in a language other than English was provided
(e.g., ``\texttt{agua}'', \emph{n} = 51), in which participants did not
provide a whole word (e.g., ``\texttt{f}'', \emph{n} = 5), and in which
participants added comments to the experimenter (e.g.,
``\texttt{unsure}'', \emph{n} = 13). In addition, we excluded data from
participants that self-rated their oral and/or written skills in Catalan
and Spanish, or any other second language as four or higher in a
five-point scale (\emph{n} = 2), were diagnosed with a language
(\emph{n} = 1), or did not contribute more than 80\% of valid trials
(\emph{n} = 9).

@@

Table~\ref{tbl-dataset} shows a summary of participants' accuracy across
Experiments 1, 2, and 3. MCMC chains in the model showed strong evidence
of convergence (\(\hat{R}<1.01\)) (see Appendix 2 for more detailed
model diagnostics). Participants translating Catalan words and
participants translating Spanish words performed equivalently, as
indicated by the regression coefficient of \emph{Group} (\(\beta\) =
-0.199, 95\% CrI = {[}-0.512, 0.129{]}, \emph{p}(ROPE) = 0.238).
Overall, participants responded less accurately to words with more CLPNs
than to words with fewer CLPNs, regardless of the amount of phonological
similarity between the presented word and its translation. This is
indicated by the size of the regression coefficient of the two-way
interaction between \emph{Similarity} and \emph{CLPN} (\(\beta\) =
-0.653, 95\% CrI = {[}-0.973, -0.313{]}, \emph{p}(ROPE) = 0). As
anticipated, participants' performance benefited from an increase in
\emph{Similarity} (\(\beta\) = 7.133, 95\% CrI = {[}6.606, 7.735{]},
\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
effect (\(\beta\) = 0.009, 95\% CrI = {[}-0.11, 0.111{]}, \emph{p}(ROPE)
= 0.925). Figure~\ref{fig-epreds-1} illustrates the posterior of the
average predictions of the model for words with different values of
\emph{Similarity} and \emph{CLPN}.

@@

We collected data for a total of 5,412 trials completed by 33 unique
participants. We excluded trials in which participants did not enter any
text (\emph{n} = 44), in which a response in a language other than
Spanish was provided (\emph{n} = 51), in which participants did not
provide a whole word (\emph{n} = 7), and in which participants added
comments to the experimenter (\emph{n} = 1). In addition, we excluded
data from participants that self-rated their oral and/or written skills
in Catalan and Spanish, or any other second language as four or higher
in a five-point scale (\emph{n} = 22), were diagnosed with a
developmental language disorder (\emph{n} = 1), or did not contribute
more than 80\% of valid trials (\emph{n} = 9). After applying
trial-level and participant-level inclusion criteria, the resulting
dataset included 1,662 trials provided by 42 unique participants. Of
those trials, 1,662 were provided by 21 unique participants who listened
to Catalan words, and 1,662 trials were provided by 21 unique
participants who listened to Spanish words. Responses given by
participants were 5.6 characters long on average (\emph{Median} = 5,
\emph{SD} = 1.6, Range = 2-12).

@@

\begin{table}

@@

\centering
\begin{tblr}[         %% tabularray outer open
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]},
cell{6}{1}={c=3}{},cell{23}{1}={c=3}{},cell{29}{1}={c=3}{},
cell{2}{1}={preto={\hspace{1em}}},
cell{3}{1}={preto={\hspace{1em}}},
cell{4}{1}={preto={\hspace{1em}}},
cell{5}{1}={preto={\hspace{1em}}},
cell{7}{1}={preto={\hspace{1em}}},
cell{8}{1}={preto={\hspace{1em}}},
cell{9}{1}={preto={\hspace{1em}}},
cell{10}{1}={preto={\hspace{1em}}},
cell{11}{1}={preto={\hspace{1em}}},
cell{12}{1}={preto={\hspace{1em}}},
cell{13}{1}={preto={\hspace{1em}}},
cell{14}{1}={preto={\hspace{1em}}},
cell{15}{1}={preto={\hspace{1em}}},
cell{16}{1}={preto={\hspace{1em}}},
cell{17}{1}={preto={\hspace{1em}}},
cell{18}{1}={preto={\hspace{1em}}},
cell{19}{1}={preto={\hspace{1em}}},
cell{20}{1}={preto={\hspace{1em}}},
cell{21}{1}={preto={\hspace{1em}}},
cell{22}{1}={preto={\hspace{1em}}},
cell{24}{1}={preto={\hspace{1em}}},
cell{25}{1}={preto={\hspace{1em}}},
cell{26}{1}={preto={\hspace{1em}}},
cell{27}{1}={preto={\hspace{1em}}},
cell{28}{1}={preto={\hspace{1em}}},
cell{30}{1}={preto={\hspace{1em}}},
cell{31}{1}={preto={\hspace{1em}}},
}                     %% tabularray inner close
\toprule
Translation & Accuracy (%) & SE \\ \midrule %% TinyTableHeader
cavall /kəβaʎ/ - horse /hɔs/             & 17.14 & 6.37 \\
llibre /ʎiβɾə/ - book /bʊk/              & 17.14 & 6.37 \\
camisa /kəmizə/ - shirt /ʃɜt/            & 16.67 & 6.21 \\
poma /poma/ - apple /æpl/                & 16.67 & 6.21 \\
Experiment 1 (cat-ENG) && \\
cama /kamə/ - leg /lɛg/                  & 11.11 & 5.24 \\
pantalón /paŋtalon/ - trousers /traʊzəz/ & 77.42 & 7.51 \\
naranja /naɾaŋxa/ - orange /ɒrɪnʤ/       & 41.94 & 8.86 \\
leche /leʧe/ - milk /mɪlk/               & 35.48 & 8.59 \\
toro /toɾo/ - bull /bʊl/                 & 33.33 & 8.61 \\
libro /liβɾo/ - book /bʊk/               & 30.00 & 8.37 \\
cebra /θebɾa/ - zebra /zibrə/            & 29.03 & 8.15 \\
pan /pan/ - bread /brɛd/                 & 29.03 & 8.15 \\
pollo /poʎo/ - chicken /ʧɪkɪn/           & 26.67 & 8.07 \\
jirafa /xiɾafa/ - giraffe /ʤɪrɑf/        & 20.69 & 7.52 \\
perro /pero/ - dog /dɒɡ/                 & 16.13 & 6.61 \\
pluma /pluma/ - feather /fɛðə/           & 16.13 & 6.61 \\
puerta /pwerta/ - door /dɔ/              & 16.13 & 6.61 \\
pie /pje/ - foot /fʊt/                   & 12.90 & 6.02 \\
caballo /kaβaʎo/ - horse /hɔs/           & 10.34 & 5.66 \\
bocadillo /bokadiʎo/ - sandwich /sænwɪʤ/ & 10.00 & 5.48 \\
Experiment 2 (spa-ENG) && \\
globo /ɡloβo/ - balloon /bəlun/          & 10.00 & 5.48 \\
fulla /fuʎə/ - hoja /oxa/                & 30.43 & 9.59 \\
ull /uʎ/ - ojo /oxo/                     & 21.74 & 8.60 \\
got /gɔt/ - vaso /baso/                  & 20.00 & 8.00 \\
entrepà /entɾəpa/ - bocadillo /bokadiʎo/ & 13.04 & 7.02 \\
Experiment 3 (cat-SPA) && \\
mirall /miɾaʎ/ - espejo /espexo/         & 12.50 & 6.75 \\
\bottomrule
\end{tblr}

\end{table}%

@@

@@

We collected data for a total of 6,016 trials completed by 64 unique
participants. Of those trials, 3,264 were provided by 32 unique
participants who listened to Catalan words, and 2,752 trials were
provided by 32 unique participants who listened to Spanish words. We
excluded trials in which participants did not enter any text (\emph{n} =
0), in which a response in a language other than English was provided
(\emph{n} = 9), in which participants did not provide a whole word
(\emph{n} = 0), and in which participants added comments to the
experimenter (\emph{n} = 0). In addition, we excluded data from
participants that self-rated their oral and/or written skills in Catalan
and Spanish, or any other second language as four or higher in a
five-point scale (\emph{n} = 2), were diagnosed with a language
(\emph{n} = 0), or did not contribute more than 80\% of valid trials
(\emph{n} = 1).

@@

MCMC chains in the model showed strong evidence of convergence
(\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
Overall, participants reported prior knowledge more often for that
Spanish words with unexpectedly high accuracy (see Discussion in
Experiment 2) than for words with expected accuracy (see
\textbf{?@fig-knowledge}). Participants reported prior knowledge of
Catalan words with unexpected accuracy as often as those with expected
accuracy. This suggests that participants in Experiment 1 may have
relied, to some extent, on their prior knowledge about form-meaning
mappings to correctly translate some Spanish words. To isolate such an
effect of prior Spanish knowledge, we run the same analysis as in
Experiment 1 on the newly collected translations from Experiment 3, now
excluding responses to words in which participants reported prior
knowledge.

@@

@@

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-baese2020perception}
Baese-Berk, M. M., McLaughlin, D. J., \& McGowan, K. B. (2020).
Perception of non-native speech. \emph{Language and Linguistics
Compass}, \emph{14}(7), e12375.

@@

\bibitem[\citeproctext]{ref-cervantes}
Cervantes, M. de, \& Grossman, E. (2002). First part of the ingenious
nobleman don quixote of la mancha. \emph{Conjunctions}, \emph{38},
207--225. \url{http://www.jstor.org/stable/24518146}

@@

\bibitem[\citeproctext]{ref-dembeck2015oberflachenubersetzung}
Dembeck, T. (2015). Oberfl{ä}chen{ü}bersetzung: The poetics and cultural
politics of homophonic translation. \emph{Critical Multilingualism
Studies}, \emph{3}(1), 7--25.

@@

\bibitem[\citeproctext]{ref-dijkstra2002architecture}
Dijkstra, T., \& Van Heuven, W. J. (2002). The architecture of the
bilingual word recognition system: {From} identification to decision.
\emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
